\section{Discussion}\label{discussion}

The 1997 measles outbreak in S\~{a}o Paulo was unexpected, having followed several years of high routine vaccination coverage, SIAs, and relatively low incidence. Further, the age distribution of cases, with a secondary mode among adults, had not been observed in previous outbreaks. Thus, while historical precedent often serves as a guide for outbreak response, in this case, such precedent would have greatly under-estimated outbreak risk and the vaccination targets necessary to control the outbreak. Here we have presented a novel approach for integrating real-time outbreak surveillance into the evaluation of an evolving outbreak in order to evaluate candidate response strategies. In doing so, we have developed a model for interpreting clinical measles surveillance that acknowledges that the correlation between clinical measles symptoms and lab confirmation of positive measles IgM serology is age-specific. Further, we have shown that relying on clinical confirmation alone can significantly bias inference about transmission rate ($R_e$) and the minimal vaccination targets required to stop an outbreak.

In the case of S\~{a}o Paulo in 1997, estimates of $R_e$ were similar as was the assessment of which campaign age targets were sufficient to reduce $R_e$ below 1, regardless of whether confirmation bias was corrected, when fit to data on 15 July. Thus, while our proposed model accounting for age-specific bias in serologic confirmation explicitly estimates the uncertainty in clinical diagnosis, it results in little practical difference in the interpretation of risk $R_e$ or candidate interventions on 15 July. However, using only the data available on 15 June, estimates based only on clinical confirmation data would have grossly under-estimated risk and over-estimated the benefit of a vaccination campaign targeting children below 5 years of age. 

Although outbreak risk can be evaluated \emph{a priori}, outbreaks themselves are often the first indication of the build-up of susceptibles or gaps in immunity. In 1997, the age distribution of cases in Sao Paulo indicated a dangerous gap in immunity among individuals between 15-35 years of age. The SIA conducted in 1987, targeting children below 14 years of age, would be expected to have immunized individuals below 23 years of age in 1997; those older than 23 years would have been born prior to a national immunization system in Brazil and would be expected to have experienced natural infection during their childhood. We estimated that excess susceptibles between 15-35 years of age may have accounted for 63\% of all susceptibles during the 1997 outbreak. We term these as \emph{excess} susceptibles because they are excess relative to the expected age distribution of susceptibles based on historical rates of natural infection, routine vaccination, and SIAs. We are unable to positively identify the source of these excess susceptibles; they may have been the result of over-estimating the coverage of previous vaccination programs, or migrants from low coverage or low transmission risk areas that were unlikely to be exposed to vaccination or natural infection. While the former explanation is possible, insufficient vaccination coverage would be expected to result in more circulating infection, which would still likely result in exposure to natural infection, and thus immunity, by adulthood. The latter explanation requires that individuals were recent immigrants to Sao Paulo and had not been exposed to either vaccination or natural infection as children in the region that the emigrated from. Measles rarely persists endemically in small populations below some critical community size \cite{Conlan_2007, 8974392}; thus it is possible that recent migrants from small villages might have not been exposed to natural infection.  A case-control trial conducted after the 1997 outbreak found that recent immigration to Sao Paulo was a significant risk factor for measles infection during the outbreak \citet{Camargo_2000} . Further, immigration rates into Sao Paulo in 1991 were highest among individuals between 15 and 30 years of age \cite{de_Moraes_2016}, which is consistent with the age distribution of the excess susceptibles estimated by our models. While this does not confirm that immigration or gaps in prior immunization were the source of the adult susceptibles during the 1997 outbreak, this analysis does suggest that these adult susceptibles may have played a significant role in the outbreak; absent the excess susceptibles, $R_e$ at the start of the outbreak would have been comfortably less than 1. Other recent measles outbreaks have exhibited this same age-profile, with an unexpectedly large number of adult cases (\emph{e.g.} Malawi \cite{Minetti_2013}, Mongolia [ref: website above], China \cite{Zheng_2015}). Thus, strategies for monitoring and targeting immunity gaps in adults (\emph{e.g} migrant susceptibles, in this case) may be useful in preventing future outbreaks. Moreover, outbreak response strategies should consider adult-targeted vaccination when surveillance indicates a large number of adult susceptibles.

Though our models account for the age distribution of susceptibles, we make very simplistic assumptions about age-specific transmission; namely that within age-class transmission is the same for all ages, and between age-class transmission decays exponentially with difference in ages. Considerable recent work has shown that age-specific mixing rates are likely to vary considerably and may be culturally specific \cite{Mossong_2008}. It is possible that higher contact (and thus, transmission) rates among adults means that adult susceptibles disproportionately contributed to this outbreak. Fitting an age-specific transmission matrix with differential age-specific rates would have greatly inflated the model complexity and is beyond the scope of the present work, but may be worthwhile in cases where age-specific contact network structure information is available \cite{327793}. In any case, we did not feel it appropriate to substitute an empirical contact matrix based on a different location (\emph{e.g.} \cite{Mossong_2008}), given the likelihood of significant geographic and cultural variation in patterns of transmission. But, we evaluated the sensitivity to our assumed network structure by estimating the model without the decay rate parameter, instead assuming an even simpler model of invariant between-group contact rates. Model predictions proved to not be strongly sensitive to this change in assumed structure (see Appendix B).

There are several potential explanations for the age-specific serological confirmation bias for cases with clinical symptoms. There are many etiologies that may generate fever and rash symptoms in young children \cite{Hutchins_2004,Ho_2014,GUY_2004,31c964}, which would increase the rate of false positives in children based on clinical symptoms alone. Further, young children may be more likely to be brought to clinic shortly after the onset of symptoms, when IgM titers may not yet have reached detectable levels. Regardless of the cause of this bias, the result is that assessing the overall age distribution of cases may tend to underestimate the role of adults if based on clinical confirmation alone. Further, as the age distribution of cases presenting with symptoms may change over the course of an outbreak (in Sao Paulo, there were relatively more young children near the start and end of the outbreak, and more adults in the middle), clinical cases may reflect time-varying confirmation and bias estimates of transmission rates. If true, by correcting for age-specific confirmation bias, we also necessarily correct for temporal variation in false positives. We see the effect of this in the higher estimated $R_e$ using the confirmation bias corrected model with the 15 June data (Figure 4), relative to that of the clinical confirmation model; though the model predicts fewer true measles cases (\emph{i.e}. false positives are removed), the exponential increase from the start of the outbreak to 15 June is steeper and this estimate is consistent with the higher $R_e$ estimate from the longer time series available on 15 July.

While we have demonstrated the potential for integrating real-time monitoring information to inform and improve outbreak response, the ability to provide model-based predictions is just one component for an operational system for adaptive control of outbreaks. For example, our approach does not take into account the investment of time and resources needed to prepare for ORI, such as vaccine procurement, microplanning, training, or social mobilization. A complete decision support system would fully integrate costs and constraints into the planning process so that management objectives can be adequately met. However, because we implemented our model using an extensible, open-source Bayesian modeling framework (PyMC), one can readily integrate it into a larger decision support system, as appropriate.

This work highlights the value of an integrated approach to model-based inference and prediction in supporting decision-making for the control of measles outbreaks. We used a Bayesian hierarchical model structure to integrate three critical sources of uncertainty that are common in emerging and re-emerging outbreaks: the structure and parameter values of the underlying disease dynamics model, partial observability of epidemic process due to imperfect reporting or diagnostics, and the current level of risk due to uncertainty in historical levels of vaccine coverage and efficacy. The full model is comprised of component sub-models, which are linked explicitly via conditional probability statements; this results in parameter estimates and evaluation of candidate vaccination targets that fully incorporate the uncertainty in the dependent variables, ultimately yielding predictions that more fully account for the lack of complete information at hand for decision-making. For example, the likely presence of a pool of excess susceptibles in the population was not evident based on historical immunization records, but was made apparent only when this information was integrated with outbreak monitoring data, via a realistic model of measles dynamics. From this, the model predicted that only campaigns targeting adults, as well as children, were likely to be effective in stopping the outbreak, and this was evident from data available as early as July 15. In practice, while targeted efforts to vaccinate school children and adults were made in 1997, before and after August campaign targeting children between 6m and 4 years, this analysis suggests, retrospectively, that there may have been evidence to support broader vaccination efforts targeting older age groups. Despite the apparent complexity of the integrated model, it was straightforward to implement and fit using readily-available, open-source software, which allowed us to iteratively evaluate key assumptions regarding model structure and prior information in terms of their influence on model estimates and predictions. We believe this approach holds promise as a structured decision-making tool for a variety of disease outbreak scenarios, by efficiently incorporating all available information and incorporating monitoring information to provide robust decision support under uncertainty.

